\chameleon{}: Untangling Type Errors Through Interactive Visualization and Exploration

### 2022-12-11 sections/chameleon.tex
Some changes

@@

Selecting a candidate expression card is a way for programmers to communicate to the type checker the knowledge that the other candidate expressions are in fact well-typed. Once a card is selected, the information in the conflicting types block changes according to the change of candidate expression. In the editor window, some error locations change highlight colors to reflect the updated knowledge of the program.

@@

Chameleon was originally a command-line tool developed in the early 2000s to improve type error reporting %and introduced a design for interactive type debugging 
for the Haskell programming language.
Unlike traditional type errors produced by the Glasgow Haskell Compiler (GHC)~\cite{ben_gamari_home_2022}, which uses a Hindley–Milner type inference system, Chameleon infers types using constraint solving. In Chameleon, Constraints are generated from the source code based on typing rules. In addition, each constraint is labeled with the location (line and column numbers) where it is generated. This set of constraints is consistent if the program is well-typed and inconsistent otherwise. When a type error occurs, an efficient algorithm is used to derive a minimal subset of the constraints that still contain inconsistencies. This subset is called a Minimal Unsatisfiable Subset (MUS). From this, Chameleon can report a list of locations, using the labels of constraints that are in the MUS. Stuckey et al.~\cite{stuckey_interactive_2003} showed that program locations linked to the constraints from an MUS are all relevant to the type error and must include the cause of the error.

@@

This has a few benefits. First, concrete types provide extra information to programmers. With a type of \texttt{(Float, Float)}, programmers may want to convey a point in 2d space. However, this information is not preserved in \texttt{(a, Float)}. Second, with concrete types, the debugging front end has more flexibility to change the representation of types. If the front end understands the type language in Haskell, it is possible to recognize and replace ground facts that are overlapping in both possible types and have the same effect. The other direction is impossible. \tf{This sentence is not very clear}

>>>>

\end{itemize}

@@

Selecting a candidate expression card is a way for programmers to communicate to the type checker the knowledge that the other candidate expressions are in fact well-typed. Once a card is selected, the information in the conflicting types panel changes according to the change of candidate expression. In the editor window, some error locations change highlight colors to reflect the updated knowledge of the program.

@@

Chameleon was originally a command-line tool developed in the early 2000s to improve type error reporting %and introduced a design for interactive type debugging 
for the Haskell programming language.
Unlike traditional type errors produced by the Glasgow Haskell Compiler (GHC)~\cite{ben_gamari_home_2022}, which uses a Hindley–Milner type inference system, Chameleon infers types using constraint solving. In Chameleon, Constraints are generated from the source code based on typing rules. In addition, each constraint is labeled with the location (line and column numbers) where it is generated. This set of constraints is consistent if the program is well-typed and inconsistent otherwise. When a type error occurs, an efficient algorithm is used to derive a minimal subset of the constraints that still contain inconsistencies. This subset is called a Minimal Unsatisfiable Subset (MUS). From this, Chameleon can report a list of locations, using the labels of constraints that are in the MUS. Stuckey et al.~\cite{stuckey_interactive_2003} showed that program locations linked to the constraints from an MUS are all relevant to the type error and must include the cause of the error. In this paper, we will refer to these locations as \textit{error locations}.

@@

This has a few benefits. First, concrete types provide extra information to programmers. With a type of \texttt{(Float, Float)}, programmers may want to convey a point in 2d space. However, this information is not preserved in \texttt{(a, Float)}. Second, with concrete types, the debugging front end has more flexibility to change the representation of types. If the front end
understands the type language in Haskell, it is possible to recognize and replace ground facts that are overlapping in both possible types and have the same effect. The other direction is impossible.

### 2022-12-11 sections/related-work.tex
Some changes

\subsection{Interactive Debugging}
The original Chameleon provides a certain level of ``interactive'' debugging through a command-line interface~\cite{stuckey_interactive_2003}, allowing the programmer to query different parts of the program. A recent study \cite{tsushima_type_2021} provided an interactive type debugging system that restores type information based on the output from CFT(Counter-factual typing)~\cite{chen_counter-factual_2014, chen_efficient_2020, chen_improving_2022} inference.  Both tools allow programmers to change their assumptions about which expression is ill-typed. While both tools show a strong influence on theoretical development, they were not evaluated in realistic studies with human participants. In addition, command line-based debugging tools suffer the same limitation as other line-based tools, namely, limited user interaction, limited graphic capability (ANSI text), and the requirement of remembering the command language.

>>>>

\subsection{Interactive Debugging}
The original Chameleon provides a certain level of ``interactive'' debugging through a command-line interface~\cite{stuckey_interactive_2003}, allowing the programmer to query different parts of the program. A recent study \cite{tsushima_type_2021} provided an interactive type debugging system that restores type information based on the output from CFT (Counter-factual typing)~\cite{chen_counter-factual_2014, chen_efficient_2020, chen_improving_2022} inference.  Both tools allow programmers to change their assumptions about which expression is ill-typed. While both tools show a strong influence on theoretical development, they were not evaluated in realistic studies with human participants. In addition, command line-based debugging tools suffer the same limitation as other line-based tools, namely, limited user interaction, limited graphic capability (ANSI text), and the requirement of remembering the command language.
